Measuring and optimizing machine learning interpretability

### 2019-03-18 2019-03-21-dagstat-imlm/dagstat-molnar.Rmd
almost finish first draft

---
title: "Measuring and optimizing machine learning interpretability"
author: "Christoph Molnar, Giuseppe Casalicchio, Bernd Bischl"
date: "2019-03-21"
output: 
    beamer_presentation
---

@@

@@

$$f(x)  = \overbrace{f_0}^\text{Intercept} + \overbrace{\sum_{j=1}^p f_j(x_j)}^\text{1st order effects} + \overbrace{\sum_{S \subseteq \{1,\ldots,p\},|S| \geq 2} f_{S}(x_S)}^{\text{Higher order effects}}
$$

$$f(x)  = \underbrace{f_0 + \sum_{j=1}^p \overbrace{f_j(x_j)}^\text{How complex?} + \overbrace{IA(x)}^{\text{How much interaction?}}}_{\text{How many feature used?}}
$$

@@

TODO: Use all plots

```{r}
knitr::include_graphics("../images/ale-intuition-3.jpg")
```
## Measure: Main effect complexity

TODO: Plot of single ALE plot with complexity (from paper)

Plus formula how to average those for all plots

## Measure: Interaction Strength

Measure main effect model with proportion of error explained (pre):

$$\text{Interaction Strength} =  \frac{\mathbb{E}(L(\hat{f}, f_{ALE}))}{\mathbb{E}(L(\hat{f}, c))}$$

@@

TODO: Wine plot
TODO: List of models

@@

TODO: Show Interpretability vs. Performance Tradeoff

@@

>>>>

---
title: "Measuring and optimizing machine learning interpretability"
author: "Christoph Molnar, Giuseppe Casalicchio, Bernd Bischl"
date: "2019-03-21"
output: 
  beamer_presentation:
    includes:
      in_header: "header_pagenrs.tex"
---

@@

## Functional Decomposition
$$f(x)  = \overbrace{f_0}^\text{Intercept} + \overbrace{\sum_{j=1}^p f_j(x_j)}^\text{1st order effects} + \overbrace{\sum_{S \subseteq \{1,\ldots,p\},|S| \geq 2} f_{S}(x_S)}^{\text{Higher order effects}}
$$

@@

$$f(x)  = \underbrace{f_0 + \sum_{j=1}^p \overbrace{f_j(x_j)}^\text{How complex?} + \overbrace{IA(x)}^{\text{How much interaction?}}}_{\text{How many feature used?}}$$

@@

```{r}
knitr::include_graphics("../images/ale-intuition-1.jpg")
```
## Accumulated Local Effects (ALE)

```{r}
knitr::include_graphics("../images/ale-intuition-2.jpg")
```

```{r}
knitr::include_graphics("../images/ale-intuition-3.jpg")
```

```{r}
knitr::include_graphics("../images/ale-intuition-4.jpg")
```
## Accumulated Local Effects (ALE)

```{r}
knitr::include_graphics("../images/ale-intuition-5.jpg")
```

## Main effect complexity

- Approximate ALE plot with linear segments
- Count number of non-zero coefficients
- Average over all features, weight with variance

```{r}
knitr::include_graphics("../images/c-demo-1.jpeg")
```
## Interaction Strength

Measure main effect model with proportion of error explained:

$$\text{Interaction Strength} =  \frac{\mathbb{E}(L(\hat{f}, f_0 + \sum_{j=1}^p f_j)}{\mathbb{E}(L(\hat{f}, c))}$$

@@

- Predict wine quality from phsicochemical properties
- Minimize MAE, number of features, interaction strength, main effect complexity
- Search across different model classes and hyperparameter settings

```{r, out.width = "70%"}
knitr::include_graphics("../images/wine-features.jpg")
```

@@

```{r}
knitr::include_graphics("../images/perf-interpret-tradeoff-1.jpeg")
```

@@

# The End
